Smart Spreading Factor Assignment for LoRaWANs

### 2019-01-13 paper.tex
SVM pdr
update figures
classifier accuracy

\par The following two sections provides a background information about LoRaWAN and LoRa. Section \ref{Other Related Works} describes other related works. Section \ref{Proposed Technique} describes proposed technique for SF assignment. Simulation environments and results are shown in Section \ref{Simulation Environment} and \ref{Simulation Results}. Finally, Section \ref{Conclusion} concludes the paper.

@@

@@

\subsection{Machine Learning Model}
\par We have been able to generate mass amount of LoRaWAN transmission logs for different topologies using the simulator. To come up with a solution for SF assignment issue, a machine learning prediction scheme is integrated to the simulator. A classifier is trained from transmission logs and this classifier is used for selecting best possible SF for the nodes. Simulator first runs a simulation with assigning random SF to nodes. Simulation results are combined into four data columns which are; X and Y coordinates of the transmission source, SF of the transmission and result of the transmission. Transmission result can be successful, interfered or under sensitivity. These data are fed to Python scikit-learn decision tree classifier. After the classifier model is generated, a second simulation is run with classifier. Simulator selects the lowest possible SF considering prediction of transmission. If a transmission of a node with lowest possible SF is predicted as interfered transmission by classifier, then simulator assigns a higher SF. If no SF is predicted as successful, then simulator again assigns lowest possible SF.

@@

\par In Figure \ref{fig:predict_pdr}, PDR plot of machine learning prediction scheme is shown. Randomly generated network topology radius is set to 5000 meters, number of GW is set to 3 and packet generation rate is set to 0.01 packet per second. Prediction model needs nodes location so three GWs are used for prediction SF scheme since three GWs are enough to roughly locate nodes positions by triangulation. Prediction scheme gives better PDR when number of nodes increases since number of interferences also increases. Prediction scheme can only improve network performance when LoRa interference is present. Also, prediction scheme gives better results when nodes are deployed close to the gateway, since nodes have margin to increase their SFs. If a node is far away from the GW, then prediction scheme cannot increase the SF to avoid interference since the SF is already high.

\begin{figure}
\centering
\includegraphics[width=\linewidth]{{predict_pdr_r5000_g3_p0.01_s3600}.png}
\caption{PDR of SF prediction scheme.}
\label{fig:predict_pdr}
\end{figure}

@@

\par As for future works, transmit power optimization can be included to proposed machine learning scheme. In this paper, we assume that nodes always use maximum transmit power for uplink transmission, however nodes close to the GW can decrease transmit power to save energy. This will make transmissions to more vulnerable to interference thus requires extra care. Optimizing another node parameter alongside SF can be interesting research. Also, other machine learning methods can be investigated for SF optimization. Reinforcement learning is a good candidate for this network optimization issue.

>>>>

\par The following two sections provides a background information about LoRa and LoRaWAN. Section \ref{Other Related Works} describes other related works. Section \ref{Proposed Technique} describes proposed technique for SF assignment. Simulation environments and results are shown in Section \ref{Simulation Environment} and \ref{Simulation Results}. Finally, Section \ref{Conclusion} concludes the paper.

@@

\par We have been able to generate mass amount of LoRaWAN transmission logs for different topologies using the simulator. To come up with a solution for SF assignment issue, a machine learning prediction scheme is integrated to the simulator. A classifier is trained from transmission logs and this classifier is used for selecting best possible SF for the nodes. Simulator first runs a simulation with assigning random SF to nodes. Simulation results are combined into four data columns which are; X and Y coordinates of the transmission source, SF of the transmission and result of the transmission. Transmission result can be successful, interfered or under sensitivity. These data are fed to Python scikit-learn decision tree classifier. After the classifier model is generated, a second simulation is run with classifier. Simulator selects the lowest possible SF considering prediction of transmission. If a transmission of a node with lowest possible SF is predicted as interfered transmission by classifier, then simulator assigns a higher SF. If no SF is predicted as successful, then simulator again assigns lowest possible SF.

@@

@@

\par In Figure \ref{fig:prediction_pdr}, PDR plot of machine learning prediction scheme is shown. Randomly generated network topology radius is set to 5000 meters, number of GW is set to 3 and packet generation rate is set to 0.01 packet per second. Prediction model needs nodes location so three GWs are used for prediction SF scheme since three GWs are enough to roughly locate nodes positions by triangulation. Prediction scheme gives better PDR when number of nodes increases since number of interferences also increases. Prediction scheme can only improve network performance when LoRa interference is present. Also, prediction scheme gives better results when nodes are deployed close to the gateway, since nodes have margin to increase their SFs. If a node is far away from the GW, then prediction scheme cannot increase the SF to avoid interference since the SF is already high.

\begin{figure}
\centering
\includegraphics[width=\linewidth]{{prediction_pdr_r5000_g3_p0.01_s3600}.png}
\caption{PDR of SF prediction scheme.}
\label{fig:prediction_pdr}
\end{figure}

\begin{table}
\centering
\caption{PDR of lowest SF selection and different SF classifier schemes.}
\label{table:prediction_pdr}
\begin{tabular}{|c|c|c|c|c|c|}
\hline
\multicolumn{3}{|c|}{\multirow{2}{*}{}}                            & \multicolumn{3}{c|}{\textbf{Number of Nodes}} \\ \cline{4-6}
\multicolumn{3}{|c|}{}                                             & 100           & 500           & 1000          \\ \hline
\multirow{12}{*}{\textbf{R (m)}} & \multirow{3}{*}{3000}  & Lowest & 97.8          & 86.0          & 72.3          \\ \cline{3-6}
                                 &                        & SVM    & 98.0          & 88.2          & 75.2          \\ \cline{3-6}
                                 &                        & DTC    & 97.8          & 89.8          & 78.7          \\ \cline{2-6}

                                 & \multirow{3}{*}{5000}  & Lowest & 96.8          & 85.5          & 71.2          \\ \cline{3-6}
                                 &                        & SVM    & 98.0          & 87.8          & 74.8          \\ \cline{3-6}
                                 &                        & DTC    & 97.7          & 90.2          & 79.8          \\ \cline{2-6}

                                 & \multirow{3}{*}{7000}  & Lowest & 97.2          & 87.5          & 76.8          \\ \cline{3-6}
                                 &                        & SVM    & 98.2          & 88.8          & 78.6          \\ \cline{3-6}
                                 &                        & DTC    & 97.8          & 90.7          & 81.6          \\ \cline{2-6}

                                 & \multirow{3}{*}{10000} & Lowest & 98.2          & 90.3          & 81.5          \\ \cline{3-6}
                                 &                        & SVM    & 98.3          & 90.3          & 81.9          \\ \cline{3-6}
                                 &                        & DTC    & 98.3          & 90.6          & 81.9          \\ \hline
\end{tabular}
\end{table}

\begin{table}
\centering
\caption{Prediction accuracy (\%) of different classifiers.}
\label{table:prediction_accuracy}
\begin{tabular}{|c|c|c|c|c|c|}
\hline
\multicolumn{3}{|c|}{\multirow{2}{*}{}}                        & \multicolumn{3}{c|}{\textbf{Number of Nodes}} \\ \cline{4-6}
\multicolumn{3}{|c|}{}                                         & 100           & 500           & 1000          \\ \hline
\multirow{8}{*}{\textbf{R (m)}} & \multirow{2}{*}{3000}  & SVM & 82.4          & 70.4          & 71.7          \\ \cline{3-6}
                                &                        & DTC & 86.0          & 67.3          & 70.4          \\ \cline{2-6}

                                & \multirow{2}{*}{5000}  & SVM & 79.5          & 69.0          & 71.1          \\ \cline{3-6}
                                &                        & DTC & 84.5          & 67.3          & 69.5          \\ \cline{2-6}

                                & \multirow{2}{*}{7000}  & SVM & 79.5          & 70.6          & 71.2          \\ \cline{3-6}
                                &                        & DTC & 84.5          & 67.7          & 69.2          \\ \cline{2-6}

                                & \multirow{2}{*}{10000} & SVM & 79.2          & 74.4          & 76.1          \\ \cline{3-6}
                                &                        & DTC & 83.8          & 70.7          & 74.3          \\ \hline
\end{tabular}
\end{table}

@@

\par As for future works, transmit power optimization can be included to proposed machine learning scheme. In this paper, we assume that nodes always use maximum transmit power for uplink transmission, however nodes close to the GW can decrease transmit power to save energy. This will make transmissions to more vulnerable to interference thus requires extra care. Also, other machine learning methods can be investigated for SF optimization. Reinforcement learning is a good candidate for this network optimization issue. TODO, node id without location TODO
